Estimating duration using survival analysis

### 2024-02-20 cis-perfect-testing.tex
ch5 clarity

\todo[inline]{make sure following paragraph makes sense in context (copy/pasted here)}
In the rest of this and the following chapter, I use survival analysis rather than the biomarker approach taken in \cref{E-ATACCC}.
This choice is motivated by computational efficiency and limitations of the CIS's study design.
The biomarker approach is computational challenging because the number of random effects scales with the number of detected infection episodes.
There are many detected episodes in CIS; in the period I will consider, around 50 times more than in ATACCC.
Therefore, the biomarker approach would be computationally challenging.
The computational challenge is compounded by the data being stored in a TRE, which has limited computational power and software.
The issue with random effects is made worse by the CIS's study design.
Infrequent testing means that most detected episodes have only a single positive test.
Hence, there is no information on the viral load trajectory and the random effects are unidentifiable.

@@

The CIS tests each individual infrequently, either weekly or monthly (see\todo{refer to section with CIS intro}).
The infrequent testing complicates analysis of the data in two ways.
First, infection episodes can go undetected.
Second, the infection episode start and end times are known imprecisely.
This section explains these issues. The rest of the chapter develops a method to tackle  them.

@@

The first complication in the analysis is that infection episodes can be undetected.
Undetected infection episodes are episodes with no associated positive tests.
Therefore, there is no information directly observed about the episode; even the number that occur is unobserved.
They occur when an episode starts and ends between tests (see \cref{perf-test:fig:truncation}).
Undetected episodes are important because they are shorter than detected episodes.
To see why this is the case, consider the undetected infection in the bottom of \cref{perf-test:fig:truncation}.
The infection episode depicted begins at time 33 and has a 10-day duration.
If the duration had been 24 days or more, the individual would still be detectable at their next test and detected then.
CIS has many undetected infections.
This is because the interval between tests in the same individual (up to 28 days, see\todo{ref section on the survey design}) is longer than most infections (median 16 days, see \cref{E-ATACCC:sec:results}).
\begin{figure}
  \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{cis-perfect-testing/truncation}}
  \caption[Undetected episodes in CIS data]{An unknown number of infection episodes in CIS are undetected. Consider the two individuals shown here: we collect identical data for them (a series of negative tests) yet the top individual was never infected and the bottom individual was. See the main text for why this is an issue. \label{perf-test:fig:truncation}}
\end{figure}

\subsection{Double interval censoring}
\todo{using double interval censored before its defined}

@@

When constructing the prior, two aspects need consideration.
Firstly, the model structure from \cref{E-ATACCC} leads to positive correlation in the posterior estimates of the hazard, which should be propagated into this analysis.
That is, the prior used in this analysis cannot be independent across the hazards.
Secondly, the uncertainty in the estimates from \cref{E-ATACCC} should be increased for this analysis.
The additional uncertainty is because the \cref{E-ATACCC} analysis extrapolates beyond the data using strong model assumptions (see\todo{reference discussion on extrapolation in ATACCC chapter}).
%Furthermore, there are differences in the study design and laboratories used between the two studies (see \cref{E-intro:sec:studies}) which may mean that results do not generalise between the studies.
I form the prior for the combination in two steps.

>>>>

\todo[inline]{Are the following paragraphs better here or in the intro chapter when motivating each chapter and the overall narrative of the thesis?}
In the rest of this and the following chapter, I use survival analysis rather than the biomarker approach taken in \cref{E-ATACCC}.
This choice is motivated by the computational limitations of the SRS and the CIS's study design (see \cref{E-intro:sec:cis}).
The biomarker approach is computational challenging because the number of random effects scales with the number of detected infection episodes.

There are many detected episodes in CIS; in the period I will consider, around 50 times more than in ATACCC.
Therefore, the biomarker approach would be computationally challenging.
The computational challenge is compounded by the data being stored in a TRE, which has limited computational power and software.
The issue with random effects is made worse by the CIS's study design.
Infrequent testing means that most detected episodes have only a single positive test.
Hence, there is no information on the viral load trajectory and the random effects are unidentifiable.

@@

The CIS tests each individual infrequently, either weekly or monthly (see \cref{E-intro:sec:cis}).
The infrequent testing complicates analysis of the data in two ways.
First, infection episodes can go undetected.
Second, the infection episode start and end times are known imprecisely.
This section explains these issues. The rest of the chapter develops a method to tackle  them.

@@

The first complication in the analysis is that infection episodes can be undetected.
Undetected infection episodes are episodes with no associated positive tests.
Therefore, there is no information directly observed about the episode; even the number that occur is unobserved.
They occur when an episode starts and ends between tests (see \cref{perf-test:fig:truncation}).
Undetected episodes are important because they are shorter than detected episodes.
To see why this is the case, consider the undetected infection in the bottom of \cref{perf-test:fig:truncation}.
The infection episode depicted begins at time 33 and has a 10-day duration.
If the duration had been 24 days or more, the individual would still be detectable at their next test and detected then.
CIS has many undetected infections.
This is because the interval between tests in the same individual (up to 28 days, see \cref{E-intro:sec:cis}) is longer than most infections (median 16 days, see \cref{E-ATACCC:sec:results}).
\begin{figure}
  \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{cis-perfect-testing/truncation}}
  \caption[Undetected episodes in CIS data]{An unknown number of infection episodes in CIS are undetected. Consider the two individuals shown here: we collect identical data for them (a series of negative tests) yet the top individual was never infected and the bottom individual was. See the main text for why this is an issue. \label{perf-test:fig:truncation}}
\end{figure}

\subsection{Unobserved episode beginning and end times}

@@

When constructing the prior, two aspects need consideration.
Firstly, the model structure from \cref{E-ATACCC} leads to positive correlation in the posterior estimates of the hazard, which should be propagated into this analysis.
That is, the prior used in this analysis cannot be independent across the hazards.
Secondly, the uncertainty in the estimates from \cref{E-ATACCC} should be increased for this analysis.
The additional uncertainty is because the \cref{E-ATACCC} analysis extrapolates beyond the data using strong model assumptions (see \cref{E-ATACCC:sec:discussion}).
%Furthermore, there are differences in the study design and laboratories used between the two studies (see \cref{E-intro:sec:studies}) which may mean that results do not generalise between the studies.
I form the prior for the combination in two steps.
